---
title: "Assignment 5"
author: "Srilaya Valmeekam"
date: "04-15-2023"
output:
  word_document: default
  pdf_document: default
---
```{r}
#Displaying the required libraries
library(cluster)
library(caret)
library(dendextend)
library(knitr)
library(factoextra)
library(readr)
```

```{r}
#Creating a data set with only numerical data by importing a dataset
library(readr)
sl_Cereals <- read.csv("~/Downloads/Cereals.csv")
Num_data <- data.frame(sl_Cereals[,4:16])
```

```{r}
#Deleting all cereals with missing values
Num_data <- na.omit(Num_data)
```

```{r}
#Normalizing data
sl_Cereals_normalise <- scale(Num_data)
```
***Apply hierarchical clustering to the adjusted data using the Euclidean distance technique***
```{r}
Dist <- dist(sl_Cereals_normalise, method = "euclidean")
H_clust <- hclust(Dist, method = "complete")
```

```{r}
#the dendogram plotting process.
plot(H_clust, cex = 0.7, hang = -1)
```


**Comparing the Clustering with single linkage, complete linkage, average linkage and Ward using the Agnes function/**
```{r}
single_Hclust <- agnes(sl_Cereals_normalise, method = "single")
complete_Hclust <- agnes(sl_Cereals_normalise, method = "complete")
average_Hclust <- agnes(sl_Cereals_normalise, method = "average")
ward_Hclust <- agnes(sl_Cereals_normalise, method = "ward")
```

***Choosing the best method***

```{r}
#Choosing the best method
print(single_Hclust$ac)
print(complete_Hclust$ac)
print(average_Hclust$ac)
print(ward_Hclust$ac)
```
#The ward strategy is the most successful one, as shown by its value of 0.9046042, which is evident given the facts provided.

***Choosing the clusters***
```{r}
pltree(ward_Hclust, cex = 0.5, hang = -1, main = "Dendrogram of agnes (Using Ward)")
rect.hclust(ward_Hclust, k = 5, border = 2:7)
S_Group <- cutree(ward_Hclust, k=5)
D_frame_2 <- as.data.frame(cbind(sl_Cereals_normalise,S_Group))
```
```{r}
fviz_cluster(list(data = D_frame_2, cluster = S_Group))
```
#From the observation mentioned above, clusters can be selected.

***Determining the stability and structure of the clusters*** 

```{r}
#Partitioning the data into A and B
set.seed(123)
partition_A <- Num_data[1:55,]
partition_B <- Num_data[56:74,]
```

```{r}
#Performing the Hierarchical Clustering of each partition while considering k = 6.
single_sl <- agnes(scale(partition_A), method = "single")
complete_sl <- agnes(scale(partition_A), method = "complete")
average_sl <- agnes(scale(partition_A), method = "average")
ward_sl <- agnes(scale(partition_A), method = "ward")
cbind(single=single_sl$ac , complete=complete_sl$ac , average= average_sl$ac , ward= ward_sl$ac)
pltree(ward_sl, cex = 0.6, hang = -1, main = "Dendogram of Agnes with Partitioned Data (Using Ward)")
rect.hclust(ward_sl, k = 6, border = 2:7)
cut_2 <- cutree(ward_sl, k = 6)
```

```{r}
#the centroids are calculated.
sl_result <- as.data.frame(cbind(partition_A, cut_2))
sl_result[sl_result$cut_2==1,]
one_centroid <- colMeans(sl_result[sl_result$cut_2==1,])
sl_result[sl_result$cut_2==2,]
two_centroid <- colMeans(sl_result[sl_result$cut_2==2,])
sl_result[sl_result$cut_2==3,]
three_centroid <- colMeans(sl_result[sl_result$cut_2==3,])
sl_result[sl_result$cut_2==4,]
four_centroid <- colMeans(sl_result[sl_result$cut_2==4,])
centroids <- rbind(one_centroid, two_centroid, three_centroid, four_centroid)
x2 <- as.data.frame(rbind(centroids[,-14], partition_B))
```

```{r}
#finding the Distance
Dist_1 <- get_dist(x2)
Matrix_1 <- as.matrix(Dist_1)
dataframe1 <- data.frame(data=seq(1,nrow(partition_B),1), Clusters =rep(0,nrow(partition_B)))
for(i in 1:nrow(partition_B)) 
  {dataframe1[i,2] <- which.min(Matrix_1[i+4, 1:4])}
dataframe1
cbind(D_frame_2$S_Group[56:74], dataframe1$Clusters)
table(D_frame_2$S_Group[56:74] == dataframe1$Clusters)
```

#Our findings from the observation mentioned above are 9 False and 10 True. As a result, we may say that the model is just partially unstable.

#The elementary public schools would like to choose a set of sl_Cereals to include in their daily cafeterias. Every day a different cereal is offered, but all sl_Cereals should support a healthy diet. For this goal, you are requested to find a cluster of “healthy Cereals''


***finding Clusters of Healthy Cereals***

```{r}
#Clustering Healthy sl_Cereals
Healthy_sl_Cereals <- sl_Cereals
Healthy_sl_Cereals_RD <- na.omit(Healthy_sl_Cereals)
clust <- cbind(Healthy_sl_Cereals_RD, S_Group)
clust[clust$S_Group==1,]
clust[clust$S_Group==2,]
clust[clust$S_Group==3,]
clust[clust$S_Group==4,]
```

```{r}
#Mean ratings are used to select the best cluster.
mean(clust[clust$S_Group==1,"rating"])
mean(clust[clust$S_Group==2,"rating"])
mean(clust[clust$S_Group==3,"rating"])
mean(clust[clust$S_Group==4,"rating"])
```

#Cluster 1 may be chosen based on the data mentioned above because it is the highest.
#Therefore, Group 1 may be considered of as the cluster for a healthy diet.

